---
title: "Dataset Setup and Usage"
author: "Oliver Keyes and David Robinson"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    keep_md: true
vignette: >
  %\VignetteIndexEntry{Dataset Setup and Usage}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

The `practice` package provides some datasets, particularly about CRAN, that make it easier to use. For maximum transparency, it is set up to construct these datasets using its own functions.

```{r}
library(practice)
```


### CRAN Metadata

For example, the `CRANmetadata` object contains a list of lists, each representing metadata information about a CRAN package. It ships with the `practice` package.

```{r}
str(CRANmetadata, max.level = 1, list.len = 6)
str(CRANmetadata[[1]], max.level = 2)
```

You can create or update this data yourself, using the `update_CRAN_metadata()` function:

```{r eval = FALSE}
practice:::update_CRAN_metadata(verbose = TRUE)
```

Note: as you might expect, this will take several minutes! You can always explore the pre-built `CRANmetadata` object first and update later at your convenience. If you do update the object, remember to re-install the package via, e.g. `devtools::install()`.

### Package sources

The sources of CRAN packages do *not* come included with the package, since they're ~ 5 GB and counting. However, `practice` provides a function to download them yourself:

```{r, eval = FALSE}
download_packages(get_package_names(), "cran_sources", quiet = FALSE)
```

This will download the sources of all packages into the `cran_sources` directory.

### Package practices

The `CRANpractices` dataset contains information on all CRAN packages and whether they follow software engineering practices. Once the package sources are downloaded as above, it can be created with:

```{r eval = FALSE}
practice:::update_CRAN_practices(src_dir = "cran_sources")
```

(This takes about two minutes).  The data looks like:

```{r}
library(practice)
library(dplyr)

CRANpractices
```

You can use it to answer questions about the frequency of particular practices:

```{r}
CRANpractices %>%
  count(license) %>%
  arrange(desc(n))

CRANpractices %>%
  count(vignette_format, vignette_builder) %>%
  ungroup() %>%
  arrange(desc(n))
```
