---
title: "Jenny Test Drive"
author: "Jenny Bryan"
date: "`r Sys.Date()`"
output:
  html_document:
    keep_md: TRUE
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, dpi = 300)
ggplot2::theme_set(ggplot2::theme_bw())
```

```{r load_packages}
library(practice)
library(dplyr)
library(ggplot2)
library(scales)
data("CRANpractices") # apparently this IS needed if first call is inline?
```

### Frequency tables and barcharts

Many variables take on a small number of values and we can digest their frequencies *en masse*.

There are `r nrow(CRANpractices)` packages in this dataset.

```{r munge-simple-factors}
## variables that make reasonable factors, w/ a modest number of levels
j_vars <- c('casing', 'alphanumeric', 'upstream_repo',
            'versioning', 'testing', 'roxygen', 'changelog')
j_vars <- setNames(j_vars, j_vars)
j_ft <- lapply(j_vars, function(x) {
  CRANpractices %>% count_(vars = list(x), sort = TRUE)
})
j_ft2 <- lapply(j_ft, function(x) {
  what <- names(x)[1]
  dots <- list(what)
  x <- x %>%
    rename_(.dots = setNames(dots, "val")) %>% 
    mutate(val = as.character(val),
           var = what,
           bar = paste(var, val, sep = ".")) %>% 
    select(bar, var, val, n)
})
j_ft3 <- bind_rows(j_ft2)
j_ft3$bar <- factor(j_ft3$bar, rev(j_ft3$bar))
```

```{r barchart-simple-factors, fig.height=6}
p <- ggplot(j_ft3, aes(x = n, y = bar))
p + geom_segment(aes(yend = bar), xend = 0, lwd = 3, colour = "grey50") +
  geom_text(aes(label = n), hjust = -0.1, size = 4) +
  facet_grid(var ~ ., scales = "free_y", space = "free_y") +
  theme(strip.background = element_blank(), strip.text = element_blank(),
        axis.title.y = element_blank()) +
  expand_limits(x = 7000)
```

Variables in `CRANpractices` that are not covered above:

```{r}
setdiff(names(CRANpractices), j_vars)
```

### License

```{r license}
lic_dat <- CRANpractices %>%
  count(license, sort = TRUE) %>% 
  mutate(license = factor(license, rev(license))) %>% 
  filter(n > 4)
p <- ggplot(lic_dat, aes(x = n, y = license))
p + geom_segment(aes(yend = license), xend = 0, lwd = 3, colour = "grey50") +
  geom_text(aes(label = n), hjust = -0.1, size = 4) +
  expand_limits(x = 2800)
```

### Downloads

```{r downloads-densityplots}
p <- ggplot(CRANpractices %>% filter(!is.na(roxygen)),
            aes(x = downloads)) + scale_x_log10(breaks = 10 ^ (1:5),
                                                labels = comma)
p + geom_density()
p + geom_density(aes(fill = roxygen), alpha = 0.5)
#p + geom_density(aes(fill = testing)) + facet_grid(testing ~ .)
p + geom_density(aes(fill = testing), alpha = 0.5)
p + geom_density(aes(fill = upstream_repo), alpha = 0.5)
```

```{r downloads-stripplots}
p <- ggplot(CRANpractices %>% filter(!is.na(roxygen)),
            aes(x = downloads)) + scale_x_log10(breaks = 10 ^ (1:5),
                                                labels = comma)
p + geom_jitter(aes(y = roxygen), alpha = 0.5)
p + geom_jitter(aes(y = testing), alpha = 0.5)
p + geom_jitter(aes(y = upstream_repo), alpha = 0.5)
```

```{r}
CRANpractices %>% 
  group_by(upstream_repo) %>% 
  summarize(no_dld = sum(downloads < 1), yes_dld = sum(downloads >= 1))
```

I find it hard to believe that organic human-driven downloads would hit essentially every single package on CRAN within a month. Are there automated systems that, e.g., download CRAN in its entirety as a matter of policy? 

### Vignettes

```{r vignettes}
vig_dat <- CRANpractices %>% 
  mutate(bar = paste(vignette_format, vignette_builder, sep = "+"))
vig_dat <- vig_dat %>%
  count(bar, sort = TRUE) %>% 
  mutate(bar = factor(bar, rev(bar))) %>% 
  filter(n > 1)
p <- ggplot(vig_dat, aes(x = n, y = bar))
p + geom_segment(aes(yend = bar), xend = 0, lwd = 3, colour = "grey50") +
  geom_text(aes(label = n), hjust = -0.1, size = 4) +
  expand_limits(x = 5500) +
  theme(axis.title.y = element_blank()) 
```

### Variables I haven't looked at yet

links_to, links_from