---
title: Software Engineering Standards in the R Community
author:
  - name: Oliver Keyes
    affiliation: Wikimedia Foundation
    email:  ironholds@gmail.com
  - name: Jennifer Bryan
    affiliation: University of British Columbia
    email:  jenny@stat.ubc.ca
  - name: David Robinson
    affiliation: Princeton University
    email:  admiral.david@gmail.com
abstract: >
  Oliver's abstract for useR goes here.
preamble: >
  % Any extra latex you need in the preamble
output: rticles::rjournal_article
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, cache = TRUE)
```

```{r load_data}
library(practice)
library(plyr)
library(dplyr)
library(tidyr)
library(ggplot2)

data("CRANpractices")
```

## Introduction

Introductory section which may include references in parentheses
\citep{R}, or cite a reference such as \citet{R} in the text.

## Best Practices

A list of practices we check, like

* Tests
* Vignettes
* roxygen2

## Practices in CRAN

Some figures and analyses, like Figure \ref{fig:CRAN_vignettes}.

```{r caption_calculations, dependson = "load_data"}
use_vignettes <- sum(CRANpractices$vignette_format != "None")
use_vignettes_percent <- 100 * use_vignettes / nrow(CRANpractices)
```

\begin{figure}
```{r vignette_figure, dependson = "load_data"}
vignette_count <- CRANpractices %>%
  count(vignette_format, vignette_builder) %>%
  ungroup() %>%
  gather(metric, choice, -n) %>%
  mutate(metric = revalue(metric, c(vignette_format = "Vignette Format",
                                    vignette_builder = "Vignette Builder"))) %>%
  filter(choice != "None") %>%
  mutate(choice = reorder(choice, n, function(x) -mean(x)))

ggplot(vignette_count, aes(choice, n)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ metric, scale = "free", ncol = 2) +
  xlab("Choice") +
  ylab("Number of packages") +
  theme_bw(base_size = 10) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
\caption{Distribution of the choice of vignette builder and format, among the `r round(use_vignettes_percent, 1)`\% of CRAN packages that use vignettes. \label{fig:CRAN_vignettes}}
\end{figure}

\bibliography{RJreferences}
